---
title: "Displaying your pretty R tables in GitHub"
author: "Laura Ellis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document
---

# Rmd file set up

Ensure that you set your .Rmd `output` to be `html_document` as per below.
```
---
title: "Displaying your pretty R tables in GitHub"
author: "Laura Ellis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document
---
```
<br/>

# Data set up

Load the necessary libraries and set some color variables which will be used during the table display.

```{r echo=TRUE}
#Load the libraries, if you don't have the library installed, run the following command per library:
#install.packages("dplyr")
library(data.table)
library(dplyr)
library(formattable)
library(DT) 

customGreen0 = "#DeF7E9"
customGreen = "#71CA97"
customRed = "#ff7f7f"

```

<br/>

## Pull in your data and format it

Download the Austin indicator data set and format it for easy table display.  

```{r}

#Original data set from: https://data.austintexas.gov/City-Government/Imagine-Austin-Indicators/apwj-7zty/data

austinData= fread('https://raw.githubusercontent.com/lgellis/MiscTutorial/master/Austin/Imagine_Austin_Indicators.csv', data.table=FALSE, header = TRUE, stringsAsFactors = FALSE)

#Format the table
i1 <- austinData %>%
  filter(`Indicator Name` %in% 
           c('Prevalence of Obesity', 'Prevalence of Tobacco Use', 
             'Prevalence of Cardiovascular Disease', 'Prevalence of Diabetes')) %>%
  select(c(`Indicator Name`, `2011`, `2012`, `2013`, `2014`, `2015`, `2016`)) %>%
  mutate (Average = round(rowMeans(
    cbind(`2011`, `2012`, `2013`, `2014`, `2015`, `2016`), na.rm=T),2), 
    `Improvement` = round((`2011`-`2016`)/`2011`*100,2))


```
<br/>


# Table Displays

We are now going to call on a few sample libraries which produce beautiful and useful html tables.


<br/>
## Example 1: Beautiful table with the fortmattable package

This example comes from my blog on [how to use the formattable package.](https://www.littlemissdata.com/blog/prettytables)
```{r}

#Create a formatter to display the red/green and arrow up/down indicators.
improvement_formatter <- formatter("span", 
                                   style = x ~ style(font.weight = "bold", 
                                                     color = ifelse(x > 0, customGreen, ifelse(x < 0, customRed, "black"))), 
                                   x ~ icontext(ifelse(x>0, "arrow-up", "arrow-down"), x)
                                   )

# Display the table using the formattable package and the custom formatter, "improvement_formatter" from above. 

formattable(i1, align =c("l","c","c","c","c", "c", "c", "c", "r"), list(
  `Indicator Name` = formatter("span", style = ~ style(color = "grey",font.weight = "bold")), 
  `2011`= color_tile(customGreen, customGreen0),
  `2012`= color_tile(customGreen, customGreen0),
  `2013`= color_tile(customGreen, customGreen0),
  `2014`= color_tile(customGreen, customGreen0),
  `2015`= color_tile(customGreen, customGreen0),
  `2016`= color_tile(customGreen, customGreen0),
  `Average` = color_bar(customRed),
  `Improvement` = improvement_formatter
))

```

<br/>

## Example 2: Pretty, interactive table with the DT package

Display a beautiful and interactive with the DT package.  I love this package for very large tables, that a user may want to interact with explore in a UI for a while before analyzing.  I used this package for my very large tables in my [pelotonR package](https://lgellis.github.io/pelotonR/)

```{r}
datatable(i1, extensions = "Scroller", width = 1000, options = list(scrollY = 200, scroller = TRUE, scrollX = 200, pageLength = 1))
```
<br/>

## More pretty R table examples

You can find a great overview of packages to use for beautiful tables in R by [David Keyes](https://twitter.com/dgkeyes) [here](https://rfortherestofus.com/2019/11/how-to-make-beautiful-tables-in-r/).

<br/>

# Export and display in GitHub

## Get the HMTL file prepared
 - Ensure that your file is called `index.Rmd`
 - Select "Knit" in the top menu bar within R Studio. This will create your file `index.htm.` 
 - Rename the HTML file to index.html
 
## Get your GitHub repo prepared
 
 - Create a new GitHub repo 
 - Go to the `Code` area > click `Add file`.
 - Add `index.html` and `index.Rmd` file from the above step
 - Navigate to the repo > `Settings` > Scroll to `GitHub Pages`
 - Select a branch.  If you are unsure, just select `Main` abd `Save`.
 - Select a theme, I selected `Minimal`.  If you need to save the new README.md file for this, do please do so.  
 - Navigate back to `Settings` > Scroll to `GitHub Pages`.  Your URL should now be displaying. It will display text similar to: `Your site is published at {URL}`
 - Celebrate because your beautiful data tables are now displayed in GitHub! 
 - Optional: Change your `README.md` file to link to your html page.


